---
format:
  pdf:
    number-sections: true
    # highlight-style: pygments
    highlight-style: files/mcshanepdf.theme
    fontsize: 11pt
    fig-pos: 'H'
    block-headings: false
    colorlinks: true
    shift-heading-level-by: 0
    geometry: 
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
    include-in-header:
      text: |
        \usepackage{fancyhdr, bm}
        \addtokomafont{disposition}{\rmfamily}
        \setcounter{section}{4}
        \lhead{ STAT2/37815 \\ Dr. McShane }
        \rhead{ YOUR NAME HERE \\ Due May 9th at 5pm }
        \chead{\textbf{\Large Homework \# 4}}
        \definecolor{codegray}{HTML}{f9f9f9}
        \definecolor{codeletter}{HTML}{002c6b}
        \let\textttOrig\texttt
        \renewcommand{\texttt}[1]{\textttOrig{\textbf{\textcolor{codeletter}{\colorbox{codegray}{#1}}}}}
        \definecolor{indigo}{RGB}{75, 0, 130}
        \definecolor{darkgreen}{RGB}{0, 128, 0}
crossref: 
  sec-prefix: Problem   # default is Section
---

\thispagestyle{fancy}

```{r setup, include = FALSE}
# Load necessary packages
# library(kableExtra)
library(formatR)

# Quarto options
knitr::opts_chunk$set(
  tidy = FALSE,     # display code as typed
  size = "small",    # slightly smaller font for code
  fig.width = 4,
  fig.height = 2.5,
  fig.align = "center", 
  cache = TRUE
)
ggplot2::theme_set(ggplot2::theme_bw())
 
# Makes kable table nice just by adding `|> kable()` to a table
# Use `escape = FALSE` to use LaTeX inside of a table
# kable = function(x, booktabs = TRUE, align = "c", format, digits = getOption("digits"), row.names = NA, col.names = NA, caption = NULL, label = NULL, format.args = list(), escape = TRUE, full_width = NULL, bootstrap_options = "basic", position = "center", latex_options = c("HOLD_position", "repeat_headers"), font_size = NULL, row_label_position = "l", ...) {
#   knitr::kable(x, booktabs = booktabs, align = align, format = format, digits = digits, row.names = row.names, col.names = col.names, caption = caption, label = label, format.args = format.args, escape = escape, ...) |>
#     kableExtra::kable_styling(full_width = full_width, bootstrap_options = bootstrap_options, position = position, latex_options = latex_options, font_size = font_size, row_label_position = row_label_position, ...)
# }
```

```{r DrM_Funks, echo = FALSE}
# Dr. McShane's functions
# ...
```


\newcommand{\ansbegin}{ \color{blue}\smallbreak\vspace{-8pt}\hrulefill \smallbreak\noindent}
\newcommand{\ansend}{\smallbreak\vspace{-8pt}\hrulefill \smallbreak\vspace{-8pt} \color{black} }

\Large \textbf{FOR THE ENTIRETY OF THIS ASSIGNMENT, DO NOT LOAD A PACKAGE.}  \normalsize _These are the only packages you should need to get started:_.

```{r libs, message = FALSE}
library(tidyverse)
library(dm)
library(DiagrammeR)
library(RSQLite)
library(RMariaDB)
library(duckdb)
library(duckplyr)
library(progress)
library(pixarfilms)
library(nycflights13)
library(parquetize)
```


\dotfill


<!-- NEW PROBLEM ----------------------------------------------->


# Question 10

## Setup

```{r }
# attach relevant packages
library(DBI)

### First steps ################################################################

# Connection -------------------------------------------------------------------

con <- dbConnect(duckdb::duckdb())
con

# Discover tables --------------------------------------------------------------

dbListTables(con)

# Populate database (normally done by other people) ---------------------------

# Magic: import tables into the database
dm::copy_dm_to(
  con,
  dm::dm_pixarfilms(),
  set_key_constraints = FALSE,
  temporary = FALSE
)

# Discover tables --------------------------------------------------------------

dbListTables(con)
dbListFields(con, "box_office")
```


## Exercises

```{r }
con
```

### 1. List all columns from the `pixar_films` table.

```{r}
dbListFields(con, "pixar_films")
```


### 2. Review the help for `dbListFields()` and `dbListTables()` and the index on <https://dbi.r-dbi.org/reference/>.

```{r}
help("dbListFields")
help("dbListTables")
browseURL("https://dbi.r-dbi.org/reference/")
```



## Question 11

### 

```{r }
# Discover tables --------------------------------------------------------------

dbListTables(con)
dbListFields(con, "pixar_films")
dbListFields(con, "academy")

# Read table -------------------------------------------------------------------

df_pixar_films <- dbReadTable(con, "pixar_films")
df_pixar_films
as_tibble(df_pixar_films)

# Execute queries --------------------------------------------------------------

dbGetQuery(con, "SELECT * FROM pixar_films")

# Assign SQL queries to character strings
sql <- "SELECT * FROM pixar_films WHERE release_date >= '2020-01-01'"

# new in R 4.1: r"()" syntax
# Kirill has used "" to indicate column names and '' for character strings
# sql <- r"(SELECT * FROM "pixar_films" WHERE "release_date" >= '2020-01-01')"
dbGetQuery(con, sql)

# Further pointers -------------------------------------------------------------

# Quoting identifiers
dbQuoteIdentifier(con, "academy")
dbQuoteIdentifier(con, "from")

# Quoting literals
dbQuoteLiteral(con, "Toy Story")
dbQuoteLiteral(con, as.Date("2020-01-01"))

# Paste queries with glue_sql()

# Parameterized queries
sql <- "SELECT count(*) FROM pixar_films WHERE release_date >= ?"
dbGetQuery(con, sql, params = list(as.Date("2020-01-01")))


# Incomplete sql query
# sql <- paste0(
#   "SELECT * FROM",
#   dbQuoteIdentifier(con, "academy"), 
#   " ", 
#   "pixar_films WHERE release_date >= ?"
# )
# 
# dbGetQuery(
#   con, 
#   sql, 
#   params = list(
#     c("Won", "Won"), 
#     c("Animated Feature", "Original Song")
#   )
# )

```


```{r }
# Reading tables: Exercises ----------------------------------------------------

con
# 1. Read the `academy` table.
# 2. Read all records from the `academy` table that correspond to awards won
#     - Hint: Use the query "SELECT * FROM academy WHERE status = 'Won'"
# 3. Use quoting and/or a query parameter to make the previous query more robust.
#     - Hint: `sql <- paste0("SELECT * FROM academy WHERE ", quoted_column, " = ?")`




```



<!-- NEW PROBLEM ----------------------------------------------->
